---
title: "Mini Data-Analysis Deliverable 1"
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr)
    package by typing the following into your **R terminal**:

<!-- -->

```         
install.packages("devtools")
devtools::install_github("UBC-MDS/datateachr")
```

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
library(ggpubr) #graphing multiple ggplot2 plots in the same figure
```

3.  Make a repository in the <https://github.com/stat545ubc-2023>
    Organization. You can do this by following the steps found on canvas
    in the entry called [MDA: Create a
    repository](https://canvas.ubc.ca/courses/126199/pages/mda-create-a-repository).
    One completed, your repository should automatically be listed as
    part of the stat545ubc-2023 Organization.

# Instructions

## For Both Milestones

-   Each milestone has explicit tasks. Tasks that are more challenging
    will often be allocated more points.

-   Each milestone will be also graded for reproducibility, cleanliness,
    and coherence of the overall Github submission.

-   While the two milestones will be submitted as independent
    deliverables, the analysis itself is a continuum - think of it as
    two chapters to a story. Each chapter, or in this case, portion of
    your analysis, should be easily followed through by someone
    unfamiliar with the content.
    [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/)
    is a good resource for what constitutes "good code". Learning good
    coding practices early in your career will save you hassle later on!

-   The milestones will be equally weighted.

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd`
file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd)
directly. Fill in the sections that are tagged with
`<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an
`.md` file by changing the YAML output settings from
`output: html_document` to `output: github_document`. Commit and push
all of your work to the mini-analysis GitHub repository you made
earlier, and tag a release on GitHub. Then, submit a link to your tagged
release on canvas.

**Points**: This milestone is worth 36 points: 30 for your analysis, and
6 for overall reproducibility, cleanliness, and coherence of the Github
submission.

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in
    tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset

The `datateachr` package by Hayley Boyce and Jordan Bourak currently
composed of 7 semi-tidy datasets for educational purposes. Here is a
brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open
    Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's
    Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning
    Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's
    Historical Hydrometric Database. It currently has 218 rows and 7
    columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open
    Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833
    rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open
    Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our
    your own *independent* data analysis. Remember to comment your code,
    be explicit about what you are doing, and write notes in this
    markdown document when you feel that context is required. As you
    advance in the project, prompts and hints to do this will be
    diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your
    goal**, or in other ways, *what you wish to achieve with this data*.
    This mini data-analysis project focuses on *data wrangling*,
    *tidying*, and *visualization*. In short, it's a way for you to get
    your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 **(1 point)** Out of the 7 datasets available in the `datateachr`
package, choose **4** that appeal to you based on their description.
Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package,
but if you have a dataset that you'd really like to use, you can include
it here. But, please check with a member of the teaching team to see
whether the dataset is of appropriate complexity. Also, include a
**brief** description of the dataset here to help the teaching team
understand your data.

<!-------------------------- Start your work below ---------------------------->

## Selected data sets to explore

1. **cancer_sample**   
2. **vancouver_trees**   
3. **steam_games**   
4. **parking_meters**

<!----------------------------------------------------------------------------->

1.2 **(6 points)** One way to narrowing down your selection is to
*explore* the datasets. Use your knowledge of dplyr to find out at least
*3* attributes about each of these datasets (an attribute is something
such as number of rows, variables, class type...). The goal here is to
have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the
cleanliness of your analysis. I added a single code chunk for you below,
but do you want to use more than one? Would you like to write more
comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->

## Exploring dimensions of each dataset

The **cancer_sample** dataset has 569 rows and 32 columns.

```{r}
#summarizing dimensions and variables of cancer_sample
glimpse(cancer_sample)
```

The **vancouver_trees** dataset has 146,611 rows and 20 columns.

```{r}
#summarizing dimensions and variables of vancouver_trees
glimpse(vancouver_trees)
```

The **steam_games** dataset has 40,833 rows and 21 columns.

```{r}
#summarizing dimensions and variables of steam_games
glimpse(steam_games)
```

The **parking_meters** dataset has 10,032 rows and 22 columns.

```{r}
#summarizing dimensions and variables of parking_meters
glimpse(parking_meters)
```

## Determining the class of each dataset

The **cancer_sample** dataset is a subclass of a tibble data frame.

```{r}
#determining class of cancer_sample
class(cancer_sample)
```

The **vancouver_trees** dataset is classed as a tibble.

```{r}
#determining class of vancouver_trees
class(vancouver_trees)
```

The **steam_games** dataset is a subclass of a tibble data frame.

```{r}
#determining class of steam_games
class(steam_games)
```

The **parking_meters** dataset is classed as a tibble.

```{r}
#determining class
class(parking_meters)
```

## Exploring variables of each dataset

The **cancer_sample** dataset consists almost entirely of numerical data with one column of character objects. There is no missing data in this data set.

```{r}
#defining class of each variable in cancer_sample
summary(cancer_sample)
```

```{r}
#previewing cancer_sample
head(cancer_sample)
```

The **vancouver_trees** dataset is equal parts between character objects and numerical data. There also appears to be a lot of missing data in the `date_planted`, `longitude` and `latitude` columns, based on the `summary()` output.

```{r}
#defining class of each variable in vancouver_trees
summary(vancouver_trees)
```

```{r}
#previewing vancouver_trees
head(vancouver_trees)
```

The **steam_games** dataset consists mostly of character objects and 4 columns of numerical data but seems difficult to parse through due to the length of some character entries. Additionally, there are several entries missing from `achievement`, `original_price` and `discount_price`.

```{r}
#defining class of each variable in steam_games
summary(steam_games)
```

```{r}
#previewing steam_games
head(steam_games)
```

The **parking_meters** dataset only contains 2 columns of numerical data. While most of the data looks numerical in nature, most columns also include units and are defined in the character class.

```{r}
#exploring classes of each variable
summary(parking_meters)
```

```{r}
#previewing parking_meters
head(parking_meters)
```

<!----------------------------------------------------------------------------->

1.3 **(1 point)** Now that you've explored the 4 datasets that you were
initially most interested in, let's narrow it down to 1. What lead you
to choose this one? Briefly explain your choice below.

<!-------------------------- Start your work below ---------------------------->

## Final Dataset Selection

The dataset I am choosing to analyze is
"**cancer_sample**". Based on my preliminary exploration of the dataset
dimensions, class and organization, I believe this dataset
will be the most straightforward to wrangle and interpret. The other
datasets I considered comprised of (1) mostly discrete, categorical
data, (2) lengthy, qualitative entries or (3) incomplete entries. These
attributes may make relationships between variables harder to parse.

<!----------------------------------------------------------------------------->

1.4 **(2 points)** Time for a final decision! Going back to the
beginning, it's important to have an *end goal* in mind. For example, if
I had chosen the `titanic` dataset for my project, I might've wanted to
explore the relationship between survival and other variables. Try to
think of 1 research question that you would want to answer with your
dataset. Note it down below.

<!-------------------------- Start your work below ---------------------------->

## Research Question

Is there a relationship between the **diagnosis**
(malignant vs. benign) and **physical attributes of the sample** e.g.
radius, texture, perimeter, area, smoothness, compactness, concavity,
symmetry, fractal dimensionality?

<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them.
Probably also a good point to grab a coffee to get ready for the fun
part!

This project is semi-guided, but meant to be *independent*. For this
reason, you will complete tasks 2 and 3 below (under the **START HERE**
mark) as if you were writing your own exploratory data analysis report,
and this guidance never existed! Feel free to add a brief introduction
section to your project, format the document with markdown syntax as you
deem appropriate, and structure the analysis as you deem appropriate. If
you feel lost, you can find a sample data analysis
[here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a
better idea. However, bear in mind that it is **just an example** and
you will not be required to have that level of complexity in your
project.

# Task 2: Exploring your dataset

If we rewind and go back to the learning objectives, you'll see that by
the end of this deliverable, you should have formulated *4* research
questions about your data that you may want to answer during your
project. However, it may be handy to do some more exploration on your
dataset of choice before creating these questions - by looking at the
data, you may get more ideas. **Before you start this task, read all
instructions carefully until you reach START HERE under Task 3**.

2.1 **(12 points)** Complete *4 out of the following 8 exercises* to
dive deeper into your data. All datasets are different and therefore,
not all of these tasks may make sense for your data - which is why you
should only answer *4*.

Make sure that you're using dplyr and ggplot2 rather than base R for
this task. Outside of this project, you may find that you prefer using
base R functions for certain tasks, and that's just fine! But part of
this project is for you to practice the tools we learned in class, which
is dplyr and ggplot2.

1.  Plot the distribution of a numeric variable.
2.  Create a new variable based on other variables in your data (only if
    it makes sense)
3.  Investigate how many missing values there are per variable. Can you
    find a way to plot this?
4.  Explore the relationship between 2 variables in a plot.
5.  Filter observations in your data according to your own criteria.
    Think of what you'd like to explore - again, if this was the
    `titanic` dataset, I may want to narrow my search down to passengers
    born in a particular year...
6.  Use a boxplot to look at the frequency of different observations
    within a single variable. You can do this for more than one variable
    if you wish!
7.  Make a new tibble with a subset of your data, with variables and
    observations that you are interested in exploring.
8.  Use a density plot to explore any of your variables (that are
    suitable for this type of plot).

2.2 **(4 points)** For each of the 4 exercises that you complete,
provide a *brief explanation* of why you chose that exercise in relation
to your data (in other words, why does it make sense to do that?), and
sufficient comments for a reader to understand your reasoning and code.

<!-------------------------- Start your work below ---------------------------->

## Exercise 4  
_Explore the relationship between 2 variables in a plot._

#### Purpose  
To explore the data, I decided to plot the distribution of
each variable and group by diagnosis. This visually can summarize which
attributes may differ between benign vs. malignant samples.

```{r}
#Storing dot plot of radius_mean vs diagnosis
radius_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = radius_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2,
              size = 0.75) +
  labs(x = "Diagnosis", y = "Radius") +
  theme(legend.position = "none")

#Storing dot plot of texture_mean vs diagnosis
texture_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = texture_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2,
              size = 0.75) + 
    labs(x = "Diagnosis", y = "Texture") +
    theme(legend.position = "none")

#Storing dot plot of perimeter_mean vs diagnosis
perimeter_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = perimeter_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2,
              size = 0.75) +
  labs(x = "Diagnosis", y = "Perimeter") +
  theme(legend.position = "none")

#Storing dot plot of area_mean vs diagnosis
area_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = area_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Area") +
  theme(legend.position = "none")

#Storing dot plot of smoothness_mean vs diagnosis
smoothness_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = smoothness_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Smoothness") +
  theme(legend.position = "none")

#Storing dot plot of compactness_mean vs diagnosis
compactness_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = compactness_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2,
              size = 0.75) +
  labs(x = "Diagnosis", y = "Compactness") +
  theme(legend.position = "none")

#Storing dot plot of concavity_mean vs diagnosis
concavity_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = concavity_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Concavity") +
  theme(legend.position = "none")

#Storing dot plot of concave_points_mean vs diagnosis
concave_points_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = concave_points_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Concave Points") +
  theme(legend.position = "none")

#Storing dot plot of symmetry_mean vs diagnosis
symmetry_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis, 
                          y = symmetry_mean)) + 
  geom_jitter(aes(colour = diagnosis),
              width = 0.2, 
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Symmetry") +
  theme(legend.position = "none")

#Storing dot plot of fractal_dimension_mean vs diagnosis
fractal_dim_plot <- ggplot(data = cancer_sample,
                      aes(x = diagnosis,
                          y = fractal_dimension_mean)) +
  geom_jitter(aes(colour = diagnosis),
              width = 0.2,
              alpha = 0.2, 
              size = 0.75) +
  labs(x = "Diagnosis", y = "Fractal Dimension") +
  theme(legend.position = "none")

#Arranging the stored dot plots into an array 
ggarrange(radius_plot, texture_plot, perimeter_plot, area_plot, #plots in first row
         smoothness_plot, compactness_plot, concavity_plot, concave_points_plot, #plots in second row
         symmetry_plot, fractal_dim_plot) #plots in third row
```

#### Conclusion    
At a glance, `radius_mean`, `texture_mean`, `perimeter_mean` and `area_mean` exhibit the strongest differential between diagnoses, with higher values in malignant samples. Other variables such as `compactness_mean`, `concavity_mean` and `concave_points_mean` illustrate a similar relationship with diagnosis, however exhibit larger spread of data. 



## Exercise 2  
_Create a new variable based on other variables in your data._

#### Purpose    
Due to the spread of data, some variables are harder to
discern if there is a clear difference based on diagnosis. For this
reason, I created a new tibble that summarizes the **average** numerical
value of **each variable**, grouped by diagnosis.

```{r}
#Create new tibble summarizing means of each variable
cancer_sample %>%
  group_by(diagnosis) %>%
  summarize(avg_radius = mean(radius_mean),
            avg_texture = mean(texture_mean),
            avg_perimeter = mean(perimeter_mean),
            avg_area = mean(area_mean), 
            avg_smoothness = mean(smoothness_mean), 
            avg_compactness = mean(compactness_mean), 
            avg_concavity = mean(concavity_mean), 
            avg_concave_points = mean(concave_points_mean), 
            avg_symmetry = mean(symmetry_mean), 
            avg_fractal_dimension = mean(fractal_dimension_mean))
```

#### Conclusion  
The average value is higher for malignant samples in all variables except `fractal_dimension_mean`. 


## Exercise 6  
_Use a boxplot to look at the frequency of different observations within a single variable._

#### Purpose  
To visualize the above table better, I generated boxplot
graphs to summarize the difference of means based on diagnosis.

```{r}
#Storing boxplot of radius_mean vs diagnosis
radius_plot <- ggplot(data = cancer_sample,
                      aes(radius_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Radius", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of texture_mean vs diagnosis
texture_plot <- ggplot(data = cancer_sample,
                      aes(texture_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Texture", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of perimeter_mean vs diagnosis
perimeter_plot <- ggplot(data = cancer_sample,
                      aes(perimeter_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Perimeter", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of area_mean vs diagnosis
area_plot <- ggplot(data = cancer_sample,
                      aes(area_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Area", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of smoothness_mean vs diagnosis
smoothness_plot <- ggplot(data = cancer_sample,
                      aes(smoothness_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Smoothness", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of compactness_mean vs diagnosis
compactness_plot <- ggplot(data = cancer_sample,
                      aes(compactness_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Compactness", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of concavity_mean vs diagnosis
concavity_plot <- ggplot(data = cancer_sample,
                      aes(concavity_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Concavity", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of concave_points_mean vs diagnosis
concave_points_plot <- ggplot(data = cancer_sample,
                      aes(concave_points_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Concave Points", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of symmetry_mean vs diagnosis
symmetry_plot <- ggplot(data = cancer_sample,
                      aes(symmetry_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Symmetry", y = "Diagnosis") +
  theme(legend.position = "none")

#Storing boxplot of fractal_dimension_mean vs diagnosis
fractal_dim_plot <- ggplot(data = cancer_sample,
                      aes(fractal_dimension_mean, diagnosis)) + 
  geom_boxplot(aes(group = diagnosis,
                   fill = diagnosis),
                   outlier.size = 0.5,
                   width = 0.5) +
  labs(x = "Fractal Dimension", y = "Diagnosis") +
  theme(legend.position = "none")

#Arranging the stored boxplots into an array
ggarrange(radius_plot, texture_plot, perimeter_plot, #plots in first row 
          area_plot, smoothness_plot, compactness_plot, #plots in second row
          concavity_plot, concave_points_plot, symmetry_plot, #plots in third row
          fractal_dim_plot, #plots in fourth row
          common.legend = TRUE, legend = "bottom") 
```

#### Conclusion    
The variables with the most discrete ranges of data are `radius_mean`, `texture_mean`, `perimeter_mean`, `area_mean`, `compactness_mean`, `concavity_mean` and `concave_points_mean`. The averages and bulk of data are numerically similar between diagnoses for `smoothness_mean`, `symmetry_mean` and `fractal_dimension_mean`.


## Exercise 8  
_Use a density plot to explore any of your variables._

#### Purpose  
Additionally, I generated density plots to better represent
the spread of data. **Note** that while the dot plots generated above (Exercise 4) also
illustrates the spread of data, the overlapping of points obscures the
density of data points near the mean. The dot plot more clearly conveys
the value and quantity of outliers, while the density plot will better
illustrate the exact values the bulk of data lies within.

```{r}
#Storing density plot of radius_mean for each diagnosis
radius_plot <- ggplot(data = cancer_sample,
                      aes(radius_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Radius", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of texture_mean for each diagnosis
texture_plot <- ggplot(data = cancer_sample,
                      aes(texture_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Texture", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of perimeter_mean for each diagnosis
perimeter_plot <- ggplot(data = cancer_sample,
                      aes(perimeter_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Perimeter", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of area_mean for each diagnosis
area_plot <- ggplot(data = cancer_sample,
                      aes(area_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Area", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of smoothness_mean for each diagnosis
smoothness_plot <- ggplot(data = cancer_sample,
                      aes(smoothness_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Smoothness", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of compactness_mean for each diagnosis
compactness_plot <- ggplot(data = cancer_sample,
                      aes(smoothness_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Compactness", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of concavity_mean for each diagnosis
concavity_plot <- ggplot(data = cancer_sample,
                      aes(concavity_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Concavity", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of concave_points_mean for each diagnosis
concave_points_plot <- ggplot(data = cancer_sample,
                      aes(concave_points_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Concave Points", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of symmetry_mean for each diagnosis
symmetry_plot <- ggplot(data = cancer_sample,
                      aes(symmetry_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
              alpha = 0.5) +
  labs(x = "Symmetry", y = "Density") +
  theme(legend.position = "none")

#Storing density plot of fractal_dimension_mean for each diagnosis
fractal_dim_plot <- ggplot(data = cancer_sample,
                      aes(fractal_dimension_mean)) + 
  geom_density(aes(group = diagnosis,
                   fill = diagnosis),
               alpha = 0.5) +
  labs(x = "Fractal Dimension", y = "Density") +
   theme(legend.position = "none")

#Arranging the stored density plots into an array
ggarrange(radius_plot, texture_plot, perimeter_plot, #plots in first row
          area_plot, smoothness_plot, compactness_plot, #plots in second row
          concavity_plot, concave_points_plot, symmetry_plot, #plots in third row
          fractal_dim_plot, #plot in fourth row
          nrow = 4, ncol = 3,
          common.legend = TRUE, legend = "bottom")
```

#### Conclusion  
Benign samples exhibit a relatively tight spread of data for all variables, while malignant samples have a much larger range of data for `radius_mean`, `perimeter_mean`, `area_mean`, `concavity_mean` and  `concave_points_mean`. 

### Summary   
* Computed averages in **Exercise 2** suggest that all variables but `fractal_dimension_mean` may have a relationship with the diagnosis, where higher values are associated with malignant samples.
* **Exercise 4** illustrated the spread of data and supported stronger correlations between high `radius_mean`, `texture_mean`, `perimeter_mean` and `area_mean` values with malignant diagnosis.
* When visualized as boxplots in **Exercise 6**, `radius_mean`, `texture_mean`, `perimeter_mean`, `area_mean`, `compactness_mean`, `concavity_mean` and `concave_points_mean` presented the most discrete data ranges between malignant and benign samples. 
* Illustrating the spread of data in **Exercise 8** revealed very large spreads of data in malignant samples for `radius_mean`, `perimeter_mean`, `area_mean`, `concavity_mean` and  `concave_points_mean`.
* With all four exercises considered, `texture_mean` and `compactness_mean` appear to correlate high values with malignancy within tighter spreads of data. However, `radius_mean`, `perimeter_mean`, `area_mean`, `concavity_mean` and `concave_points_mean` also correlate high values with malignancy and exhibit a smaller numerical range of overlap between diagnoses. 

<!----------------------------------------------------------------------------->

# Task 3: Choose research questions

**(4 points)** So far, you have chosen a dataset and gotten familiar
with it through exploring the data. You have also brainstormed one
research question that interested you (Task 1.4). Now it's time to pick
4 research questions that you would like to explore in Milestone 2!
Write the 4 questions and any additional comments below.

<!--- *****START HERE***** --->

## Further Research Questions    
In Task 2, I investigated the relationship between the **physical attributes** of cancer samples and **diagnosis**, within the dataset `cancer_samples`. In Task  3, I would like to explore these relationships further, evaluate their strength and determine the level of confidence they can predict malignancy. 

1. Which physical attribute is the strongest
indicator of diagnosis?   
2. Is there a numerical range of this attribute
that can define a benign vs. malignant tumor?   
3. Do samples with large
margins of error (SE value) still align with the above ranges of benign
vs. malignant?    
4. For discrete ranges of this attribute, what is the
likelihood that the sample is benign vs. malignant?


<!----------------------------->

# Overall reproducibility/Cleanliness/Coherence Checklist

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major
continuity errors. An example of a major continuity error is having a
data set listed for Task 3 that is not part of one of the data sets
listed in Task 1.

## Error-free code (3 points)

For full marks, all code in the document should run without error. 1
point deduction if most code runs without error, and 2 points deduction
if more than 50% of the code throws an error.

## Main README (1 point)

There should be a file named `README.md` at the top level of your
repository. Its contents should automatically appear when you visit the
repository on GitHub.

Minimum contents of the README file:

-   In a sentence or two, explains what this repository is, so that
    future-you or someone else stumbling on your repository can be
    oriented to the repository.
-   In a sentence or two (or more??), briefly explains how to engage
    with the repository. You can assume the person reading knows the
    material from STAT 545A. Basically, if a visitor to your repository
    wants to explore your project, what should they know?

Once you get in the habit of making README files, and seeing more README
files in other projects, you'll wonder how you ever got by without them!
They are tremendously helpful.

## Output (1 point)

All output is readable, recent and relevant:

-   All Rmd files have been `knit`ted to their output md files.
-   All knitted md files are viewable without errors on Github. Examples
    of errors: Missing plots, "Sorry about that, but we can't show files
    that are this big right now" messages, error messages from broken R
    code
-   All of these output files are up-to-date -- that is, they haven't
    fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were
    knitting an Rmd to html, but then changed the output to be only a
    markdown file, then the html file is a relic and should be deleted.

(0.5 point deduction if any of the above criteria are not met. 1 point
deduction if most or all of the above criteria are not met.)

Our recommendation: right before submission, delete all output files,
and re-knit each milestone's Rmd file, so that everything is up to date
and relevant. Then, after your final commit and push to Github, CHECK on
Github to make sure that everything looks the way you intended!

## Tagged release (0.5 points)

You've tagged a release for Milestone 1.

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and
Vincenzo Coia for launching.
